\singlespacing \vspace{-1.5cm} \LARGE Trade-offs among plant reproductive traits determine interactions with floral visitors

### 2021-09-16 Manuscript/Manuscript.Rmd
wording edit from reproductive to reproductive biology;
and

describing new model with prsence absence. Now I have to redo analysis of supp. mat to match all analyses

__Plant life strategies are often delimited by vegetative and physiological traits but little is known about how floral and reproductive traits drive these strategies, and in turn shape plant interactions with floral visitors. Here, we compiled 13 floral, 4 reproductive and 3 vegetative traits for 1,506 plant species from 28 plant-pollinator network studies across 18 different countries. We investigated the associations among these traits, pollinator visitation and the functional role of plant species within the networks (interaction frequency, normalized degree and specialization). We found that 51.8% of trait variation was explained by two independent axes that encompassed plant form and function. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size (PC1, 26.72%). The second axis indicated a trade-off for the level of pollinator dependency (PC2, 25.08%). Although the main axes of trait variation did not fully explain pollinator visitation rates, different plant life strategies were associated with visitation rates and pollinator functional groups. Overall, the main traits that determined plant species’ functional roles were height, nectar concentration, pollen grains per flower, number of ovules, style length, selfing level and flower width. Our results highlight the need to consider plant reproductive and floral traits to improve understanding of plant life strategies and plant-pollinator interactions at broader spatial scales.__

@@

There is an astonishing diversity of floral structures and plant reproductive strategies among flowering plants [@barrett2002; @schiestl2013], which have long been of interest to pollination biologists in terms of their relevance to plant-pollinator interactions. However, most studies that have explored reproductive (e.g., mating and compatibility systems) and floral trait (e.g., flower size or nectar provision) variation have concentrated on the individual or community level and thus, broader macroecological patterns remain poorly investigated [@carvalheiro2014; @baude2016; @munoz2016; @moeller2017; @grossenbacher2017]. Indeed, studies depicting species’ life history strategies generally focus on vegetative traits and rarely consider reproductive traits [@salguero2016, @ruger2018]. As a consequence, a unified framework that explores the compromises among floral traits and their relevance to plant life strategies is currently lacking [@roddy2021]. At the same time, there is growing interest in the determinants of plant-pollinator interactions via trait-based approaches [@fenster2004] and trait-matching analyses [@bartomeus2016]. However, floral traits have been overlooked beyond highly specialised plant-pollinator systems [@dellinger2020; @roddy2021] and the role of plant reproductive biology remains little explored in plant-pollinator interactions (but see references [@tur2013; @devaux2014]).

  With the recent availability of large trait databases, plant ecological strategies are increasingly being examined [e.g., TRY; @kattge2011 and COMPADRE; @salguero2015], and are facilitating the identification of global patterns and constraints of plant form and function [@salguero2016; @diaz2016; @carmona2021]. However, the main focus has been on vegetative traits such as leaf [@wright2004] or wood [@chave2009] trade-offs with little or no attention given to reproductive and floral traits [@evojtko2020], also critical to plant form and function. For instance, short lived versus perennial species tend to have low versus high levels of outcrossing, respectively [@barrett2003; @moeller2017]. Further, outcrossing levels are positively correlated with flower size [@goodwillie2010]. In addition, the presence of costly rewards (e.g., pollen or nectar) and showy flowers or floral displays can only be understood through consideration of plant species’ reliance upon animal pollination (pollinator dependence) and their role in attracting pollinators [@ollerton2011]. Hence, exploring plant life strategies with reproductive and floral trade-offs, in conjunction with their pollinator dependence, is necessary for a balanced understanding of plant economics.

@@

Here, we explore the potential trade-offs among plant floral and reproductive traits and how these influence the structure of plant-pollinator networks. First, we identify the major axes of floral and reproductive trait variation and trade-offs that determine plant form and function. Second, we investigate how plant species’ position in trait-space influences their interaction strength with different guilds of floral visitors. Finally, we investigate how the main axes of trait variation and individual traits influence plant species roles within networks.

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=27,fig.pos = "!H",fig.cap= "**Fig. 3 | Interaction (yes/no) and visitation rates across the three main axes of trait variation per floral visitor guild.** Fitted posterior estimates of the presence/absence of interaction (a, b and c) and number of visits (d, e and f) made by the different floral visitors guilds in relation to PC1, PC2 and PC3. PC1 represents the flower number - flower size trade-off, PC2 represents the pollinator dependence trade-off and PC3, the style length trade-off."}

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=4.25, fig.width=12,fig.pos = "!H",fig.cap= "**Fig. 4 | contribution of traits in plant's network roles.** Regression tree analysis of interaction frequency (log-transformed), normalized degree and specialization for the subset of species with quantitative data for pollen and nectar traits. The superior value inside the node indicates the mean value of the different species-level metric and the lower value, the percentage of species that are considered in each node. Thus, the top node has the mean value of the named trait for the 100% of species. Each node has a yes/no question and when the condition is fulfilled, the branch turns to the ‘yes’ direction and when not, to the ‘no’ direction. This rationale is followed in all the regression trees as indicated in the first branch division of the topmost node of each tree."}

@@

Here, we show that plant species exhibit clear trade-offs in their floral, reproductive and vegetative traits. These trade-offs are differentiated on three main axes of trait variation: (i) flower number - flower size, (ii) pollinator dependence and (iii) style length. Although these axes of trait variation did not explain overall visitation rates, we found that plant life strategies were clearly associated with different floral visitor guilds. Interestingly, pollen and nectar related traits were better than all other traits for characterizing plant species functional roles within their interaction networks.

Over 50% of plant trait variation was captured by the flower number - flower size and pollinator dependence trade-offs. Variation in these two axes of trait variation appears to be associated with the commonly described ‘fast-slow continuum’ in plant [@salguero2016] and animal [@healy2019] life-history strategies as indicated by the different floral and reproductive traits associated with plant height, life form and lifespan. The ‘slow’ part of the continuum (i.e., tall trees and shrubs) had plant species with many flowers, few ovules, higher pollinator dependence, greater occurrence of self-incompatibility and more complex breeding systems (e.g., monoecious and dioecious species). In contrast, species that employed the ‘fast’ strategy (i.e., short herbs), had fewer flowers, more ovules, greater occurrence of self compatibility and lower pollinator dependence. Further, on the first two axes of trait variation, we found additional support for the positive association already described in a previous study between outcrossing rate and floral display [@goodwillie2010]. This trend was further confirmed when looking at the specific correlation between floral display and autonomous selfing for our dataset (see Supplementary Fig. S10).

@@

To conclude, we provide the first description of plant floral and reproductive trade-offs using a large global dataset of plant traits and interactions with floral visitors. This allowed us to identify the major reproductive strategies of flowering plants and how plant life strategies determine their interaction with floral visitor guilds. Although the percentage of explained variation found with the first two axes (~50%) is lower than other studies that have evaluated life histories with vegetative traits [@diaz2016; @carmona2021], this percentage is consistent with the largest and most recent study that has characterized life strategies with both vegetative and reproductive (i.e., semelparity and iteroparity) traits [@salguero2016]. Future work needs to integrate these floral and reproductive compromises with the already described vegetative and physiological trade-offs [@diaz2016] to create a more comprehensive spectrum of trait variation [@roddy2021]. Further, the varying level of phylogenetic signal among traits deserves further attention to understand evolutionary changes on mating and flower morphology in response to pollinators [@gervasi2017; @mackin2021]. Finally, a more complete description of the macroecological patterns of plant-pollinator interactions including unrepresented areas of the world [@poisot2021] and traits that have been largely ignored (e.g., floral rewards and plant mating system) are needed to progress in our understanding of the global patterns in plant-pollinator interactions.

@@

**Functional traits.** We selected 20 different functional traits based on their relevance to plant reproduction and data availability (Table 1). These included twelve quantitative and eight categorical traits belonging to three broader trait groupings (13 floral, 4 reproductive and 3 vegetative, Supplementary Information). For each plant species, we undertook an extensive literature and online search across a wide range of resources (plant databases, online floras, books, journals and images). From a total of 30,120 cells (20 columns × 1,506 species) we were able to fill 24,341 cells (80.8% of the dataset, see Supplementary Fig. S1 for missing values information for each trait).

@@

Type <- c("Vegetative", "Floral", "Floral", "Floral", "Floral", "Floral", "Floral","Floral","Floral","Floral","Floral","Reproductive")

Type_1 <- c("Vegetative", "Vegetative", "Floral", "Floral", "Floral", "Reproductive", "Reproductive", "Reproductive","","","","")

>>>>

__Plant life strategies are often delimited by vegetative and physiological traits but little is known about how floral and reproductive biology traits drive these strategies, and in turn shape plant interactions with floral visitors. Here, we compiled 13 floral, 4 reproductive biology and 3 vegetative traits for 1,506 plant species from 28 plant-pollinator network studies across 18 different countries. We investigated the associations among these traits, pollinator visitation and the functional role of plant species within the networks (interaction frequency, normalized degree and specialization). We found that 51.8% of trait variation was explained by two independent axes that encompassed plant form and function. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size (PC1, 26.72%). The second axis indicated a trade-off for the level of pollinator dependency (PC2, 25.08%). Although the main axes of trait variation did not fully explain pollinator visitation rates, different plant life strategies were associated with visitation rates and pollinator functional groups. Overall, the main traits that determined plant species’ functional roles were height, nectar concentration, pollen grains per flower, number of ovules, style length, selfing level and flower width. Our results highlight the need to consider plant reproductive and floral traits to improve understanding of plant life strategies and plant-pollinator interactions at broader spatial scales.__

@@

There is an astonishing diversity of floral structures and plant reproductive strategies among flowering plants [@barrett2002; @schiestl2013], which have long been of interest to pollination biologists in terms of their relevance to plant-pollinator interactions. However, most studies that have explored plant reproductive biology (e.g., mating and compatibility systems) and floral trait (e.g., flower size or nectar provision) variation have concentrated on the individual or community level and thus, broader macroecological patterns remain poorly investigated [@carvalheiro2014; @baude2016; @munoz2016; @moeller2017; @grossenbacher2017]. Indeed, studies depicting species’ life history strategies generally focus on vegetative traits and rarely consider reproductive biology traits [@salguero2016, @ruger2018]. As a consequence, a unified framework that explores the compromises among floral traits and their relevance to plant life strategies is currently lacking [@roddy2021]. At the same time, there is growing interest in the determinants of plant-pollinator interactions via trait-based approaches [@fenster2004] and trait-matching analyses [@bartomeus2016]. However, floral traits have been overlooked beyond highly specialised plant-pollinator systems [@dellinger2020; @roddy2021] and the role of plant reproductive biology remains little explored in plant-pollinator interactions (but see references [@tur2013; @devaux2014]).

  With the recent availability of large trait databases, plant ecological strategies are increasingly being examined [e.g., TRY; @kattge2011 and COMPADRE; @salguero2015], and are facilitating the identification of global patterns and constraints of plant form and function [@salguero2016; @diaz2016; @carmona2021]. However, the main focus has been on vegetative traits such as leaf [@wright2004] or wood [@chave2009] trade-offs with little or no attention given to reproductive biology and floral traits [@evojtko2020], also critical to plant form and function. For instance, short lived versus perennial species tend to have low versus high levels of outcrossing, respectively [@barrett2003; @moeller2017]. Further, outcrossing levels are positively correlated with flower size [@goodwillie2010]. In addition, the presence of costly rewards (e.g., pollen or nectar) and showy flowers or floral displays can only be understood through consideration of plant species’ reliance upon animal pollination (pollinator dependence) and their role in attracting pollinators [@ollerton2011]. Hence, exploring plant life strategies with reproductive biology and floral trade-offs, in conjunction with their pollinator dependence, is necessary for a balanced understanding of plant economics.

@@

Here, we explore the potential trade-offs among plant floral and reproductive biology traits and how these influence the structure of plant-pollinator networks. First, we identify the major axes of trait variation and trade-offs that determine plant form and function. Second, we investigate how plant species’ position in trait-space influences their interaction strength with different guilds of floral visitors. Finally, we investigate how the main axes of trait variation and individual traits influence plant species roles within networks.

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=27,fig.pos = "!H",fig.cap= "**Fig. 3 | Interaction (yes/no) and visitation rates across the three main axes of trait variation per floral visitor guild.** Fitted posterior estimates of the presence/absence of interaction (a, b and c) and number of visits (d, e and f) made by the different floral visitors guilds in relation to PC1, PC2 and PC3. PC1 represents the flower number - flower size trade-off, PC2 represents the pollinator dependence trade-off and PC3, the style length trade-off. For visualization purposes, due to large differences between the visitation rates of bees and the rest of guilds, the number of visits was log-transformed (Y-axis of lower panel)."}

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=4.25, fig.width=12,fig.pos = "!H",fig.cap= "**Fig. 4 | Contribution of traits in plant's network roles.** Regression tree analysis of interaction frequency (log-transformed), normalized degree and specialization for the subset of species with quantitative data for pollen and nectar traits. The superior value inside the node indicates the mean value of the different species-level metric and the lower value, the percentage of species that are considered in each node. Thus, the top node has the mean value of the named trait for the 100% of species. Each node has a yes/no question and when the condition is fulfilled, the branch turns to the ‘yes’ direction and when not, to the ‘no’ direction. This rationale is followed in all the regression trees as indicated in the first branch division of the topmost node of each tree."}

@@

Here, we show that plant species exhibit clear trade-offs in their floral, reproductive biology and vegetative traits. These trade-offs are differentiated on three main axes of trait variation: (i) flower number - flower size, (ii) pollinator dependence and (iii) style length. Although these axes of trait variation did not explain overall visitation rates, we found that plant life strategies were clearly associated with different floral visitor guilds. Interestingly, pollen and nectar related traits were better than all other traits for characterizing plant species functional roles within their interaction networks.

Over 50% of plant trait variation was captured by the flower number - flower size and pollinator dependence trade-offs. Variation in these two axes of trait variation appears to be associated with the commonly described ‘fast-slow continuum’ in plant [@salguero2016] and animal [@healy2019] life-history strategies as indicated by the different floral and reproductive biology traits associated with plant height, life form and lifespan. The ‘slow’ part of the continuum (i.e., tall trees and shrubs) had plant species with many flowers, few ovules, higher pollinator dependence, greater occurrence of self-incompatibility and more complex breeding systems (e.g., monoecious and dioecious species). In contrast, species that employed the ‘fast’ strategy (i.e., short herbs), had fewer flowers, more ovules, greater occurrence of self compatibility and lower pollinator dependence. Further, on the first two axes of trait variation, we found additional support for the positive association already described in a previous study between outcrossing rate and floral display [@goodwillie2010]. This trend was further confirmed when looking at the specific correlation between floral display and autonomous selfing for our dataset (see Supplementary Fig. S10).

@@

To conclude, we provide the first description of plant floral and reproductive biology trade-offs using a large global dataset of plant traits and interactions with floral visitors. This allowed us to identify the major reproductive strategies of flowering plants and how plant life strategies determine their interaction with floral visitor guilds. Although the percentage of explained variation found with the first two axes (~50%) is lower than other studies that have evaluated life histories with vegetative traits [@diaz2016; @carmona2021], this percentage is consistent with the largest and most recent study that has characterized life strategies with both vegetative and reproductive (i.e., semelparity and iteroparity) traits [@salguero2016]. Future work needs to integrate these floral and reproductive compromises with the already described vegetative and physiological trade-offs [@diaz2016] to create a more comprehensive spectrum of trait variation [@roddy2021]. Further, the varying level of phylogenetic signal among traits deserves further attention to understand evolutionary changes on mating and flower morphology in response to pollinators [@gervasi2017; @mackin2021]. Finally, a more complete description of the macroecological patterns of plant-pollinator interactions including unrepresented areas of the world [@poisot2021] and traits that have been largely ignored (e.g., floral rewards and plant mating system) are needed to progress in our understanding of the global patterns in plant-pollinator interactions.

@@

**Functional traits.** We selected 20 different functional traits based on their relevance to plant reproduction and data availability (Table 1). These included twelve quantitative and eight categorical traits belonging to three broader trait groupings (13 floral, 4 reproductive biology and 3 vegetative, Supplementary Information). For each plant species, we undertook an extensive literature and online search across a wide range of resources (plant databases, online floras, books, journals and images). From a total of 30,120 cells (20 columns × 1,506 species) we were able to fill 24,341 cells (80.8% of the dataset, see Supplementary Fig. S1 for missing values information for each trait).

@@

Type <- linebreak(c("Vegetative", "Floral", "Floral", "Floral", "Floral", "Floral", "Floral","Floral","Floral","Floral","Floral","Reproductive \n biology"),align = c("c"))

Type_1 <- linebreak(c("Vegetative", "Vegetative", "Floral", "Floral", "Floral", "Reproductive \n biology", "Reproductive \n biology", "Reproductive \n biology","","","",""),align = c("c"))
